---
title: "MSDS Second Case Study"
author: "Alec Lepe, Che Cobb"
date: "11/28/2018"
output: 
  html_document:
  keep_md: true
---
### Load Libraries
```{r setup, include=FALSE}
#install.packages("data.table")
#install.packages("readxl")
#install.packages("ggplot2")
#install.packages("reshape")
library("data.table")
library("readxl")
library("ggplot2")
library("reshape")
knitr::opts_chunk$set(fig.width = 10, fig.length=10, fig.asp = 1/3,  message=FALSE)
```
### Reproducibility: Session Info
```{r aeessionInfo}
sessionInfo()
```
### Load Data
```{r LoadData}
casedata=data.table::as.data.table(readxl::read_excel("~/MSDS-6306-Second-Case-Study/Data/CaseStudy2-data.xlsx"))
save(casedata,file="~/MSDS-6306-Second-Case-Study/Data/CaseStudy2-data.RData")
load("~/MSDS-6306-Second-Case-Study/Data/CaseStudy2-data.RData")
dim(casedata)
str(casedata)
```
### Detect and Delete Non-Relevant Data
``` {r DeleteData}
cns=c()
for (cn in colnames(casedata)){
  if(length(table(casedata[[cn]]))==1){
    print(cn)
    cns=append(cn,cns)
  }
}
for (cn in cns){
  print(paste(cn,names(table(casedata[[cn]])),table(casedata[[cn]])))
  casedata=casedata[,!cn,with=F]
}
```
Deleted: *Standard Hours*, *Over18*, and *EmplyoeeCount* where there is only one value.
``` {r AttritionCount}
barplot(table(casedata[["Attrition"]]), main="Attrition Count",col=c("Green","Red"))
table(casedata[["Attrition"]])/nrow(casedata)
```
### How does age relate to attrition?
``` {r AttritionCountByAge}
ID="Age"
yes=casedata[,sum(Attrition=="Yes"),ID]; colnames(yes)[2]="Yes"
no=casedata[,sum(Attrition=="No"),ID];   colnames(no)[2]="No"
temp=merge(yes,no,all=T)
temp=melt(temp,id=ID)
temp=as.data.frame(temp)
colnames(temp)=c(ID,"Attrition","Count")
ggplot(temp,aes(fill=Attrition,y=Count,x=eval(as.symbol(ID))))+geom_bar(stat="identity")+xlab(ID)
```
Here we can see that attrition is greatest between 29 and 35.
``` {r AttritionProportionByAge}
temp=merge(temp,as.data.table(temp)[,sum(Count),ID])
colnames(temp)[ncol(temp)]="Total"
temp[["Proportion"]]=temp[["Count"]]/temp[["Total"]]
ggplot(temp,aes(fill=Attrition,y=Proportion,x=eval(as.symbol(ID))))+geom_bar(stat="identity")+xlab(ID)+ggtitle(paste(ID,"vs Attrition"))
```
Here we see that attrition is greater at earlier ages, with a noticeable peak at age 58.
``` {r age58}
temp[temp[["Age"]]==58,]
```
### How does business travel relate to attrition?
``` {r AttritionCountByBusinessTravel}
ID="BusinessTravel"
yes=casedata[,sum(Attrition=="Yes"),ID]; colnames(yes)[2]="Yes"
no=casedata[,sum(Attrition=="No"),ID];   colnames(no)[2]="No"
temp=merge(yes,no,all=T)
temp=melt(temp,id=ID)
temp=as.data.frame(temp)
colnames(temp)=c(ID,"Attrition","Count")
temp=merge(temp,as.data.table(temp)[,sum(Count),ID])
colnames(temp)[ncol(temp)]="Total"
temp[["Proportion"]]=temp[["Count"]]/temp[["Total"]]
ggplot(temp,aes(fill=Attrition,y=Proportion,x=eval(as.symbol(ID))))+geom_bar(stat="identity")+xlab(ID)+ggtitle(paste(ID,"vs Attrition"))
```
Here we can see that attrition, proportionally, is greatest for employees who travel frequently and the least for employees who do not travel.
### Automatically detecting single predictors
We can go on and on plotting individual predictors, but let's try to automate the process.
This R code will go through every numeric variable and see if they can predict attrition.
```{r determinebestvariables}
tempcasedata=casedata
tempcasedata[tempcasedata[["Attrition"]]=="Yes",][["Attrition"]]="1"
tempcasedata[tempcasedata[["Attrition"]]=="No",][["Attrition"]]="0"
mode(tempcasedata[["Attrition"]])="numeric"
pmatrix=matrix(ncol=2,nrow=0)
for (i in setdiff(colnames(tempcasedata),"Attrition")){
  if (mode(tempcasedata[[i]])=='numeric'){
  formula=paste0("Attrition~",i,sep="")
  model=lm(formula,tempcasedata)
  pmatrix=rbind(pmatrix,c(i,summary(model)$coefficients[8]))
}
}
pmatrix=as.data.table(pmatrix[order(as.numeric(pmatrix[,2])),])
colnames(pmatrix)=c("Variable","P-Value")
```
Here are the top 6 variables, based on low P-Values.
``` {r variables}
head(pmatrix)
cat(head(pmatrix)[["Variable"]])
```
### Plotting Statistically Significant Data
Excluding age as we previouslyu plotted that out.
``` {r TotalWorkingYears}
ID="TotalWorkingYears"
yes=casedata[,sum(Attrition=="Yes"),ID]; colnames(yes)[2]="Yes"
no=casedata[,sum(Attrition=="No"),ID];   colnames(no)[2]="No"
temp=merge(yes,no,all=T)
temp=melt(temp,id=ID)
temp=as.data.frame(temp)
colnames(temp)=c(ID,"Attrition","Count")
temp=merge(temp,as.data.table(temp)[,sum(Count),ID])
colnames(temp)[ncol(temp)]="Total"
temp[["Proportion"]]=temp[["Count"]]/temp[["Total"]]
ggplot(temp,aes(fill=Attrition,y=Proportion,x=eval(as.symbol(ID))))+geom_bar(stat="identity")+xlab(ID)+ggtitle(paste(ID,"vs Attrition"))
```

Here we see that attrition generally decreases as total working years increases with an outlier at age 40.
``` {r TotalWorkingYears40}
temp[temp[["TotalWorkingYears"]]==40,]
```
``` {r JobLevel}
ID="JobLevel"
yes=casedata[,sum(Attrition=="Yes"),ID]; colnames(yes)[2]="Yes"
no=casedata[,sum(Attrition=="No"),ID];   colnames(no)[2]="No"
temp=merge(yes,no,all=T)
temp=melt(temp,id=ID)
temp=as.data.frame(temp)
colnames(temp)=c(ID,"Attrition","Count")
temp=merge(temp,as.data.table(temp)[,sum(Count),ID])
colnames(temp)[ncol(temp)]="Total"
temp[["Proportion"]]=temp[["Count"]]/temp[["Total"]]
ggplot(temp,aes(fill=Attrition,y=Proportion,x=eval(as.symbol(ID))))+geom_bar(stat="identity")+xlab(ID)+ggtitle(paste(ID,"vs Attrition"))
```

Here we see attrition roughly vary between job level.
``` {r YearsInCurrentRole}
ID="YearsInCurrentRole"
yes=casedata[,sum(Attrition=="Yes"),ID]; colnames(yes)[2]="Yes"
no=casedata[,sum(Attrition=="No"),ID];   colnames(no)[2]="No"
temp=merge(yes,no,all=T)
temp=melt(temp,id=ID)
temp=as.data.frame(temp)
colnames(temp)=c(ID,"Attrition","Count")
temp=merge(temp,as.data.table(temp)[,sum(Count),ID])
colnames(temp)[ncol(temp)]="Total"
temp[["Proportion"]]=temp[["Count"]]/temp[["Total"]]
ggplot(temp,aes(fill=Attrition,y=Proportion,x=eval(as.symbol(ID))))+geom_bar(stat="identity")+xlab(ID)+ggtitle(paste(ID,"vs Attrition"))
```

Here we see attrition generally decrease as years in current role increases with seemingly an outlier at age 15.
``` {r yearsincurrentrole15}
temp[temp[["YearsInCurrentRole"]]==15,]
```

``` {r MonthlyIncome}
ID="MonthlyIncome2"
casedata[["MonthlyIncome2"]]=round(casedata[["MonthlyIncome"]]/1000)
yes=casedata[,sum(Attrition=="Yes"),ID]; colnames(yes)[2]="Yes"
no=casedata[,sum(Attrition=="No"),ID];   colnames(no)[2]="No"
temp=merge(yes,no,all=T)
temp=melt(temp,id=ID)
temp=as.data.frame(temp)
colnames(temp)=c(ID,"Attrition","Count")
temp=merge(temp,as.data.table(temp)[,sum(Count),ID])
colnames(temp)[ncol(temp)]="Total"
temp[["Proportion"]]=temp[["Count"]]/temp[["Total"]]
ggplot(temp,aes(fill=Attrition,y=Proportion,x=MonthlyIncome2))+geom_bar(stat="identity")+xlab("Rounded Monthly Income ($K)")+ggtitle("Rounded Monthly Income ($K) vs Attrition")
```

Here we see attrition generally decrease with increases in monthly income levels.
``` {r YearsWithCurrManager}
ID="YearsWithCurrManager"
yes=casedata[,sum(Attrition=="Yes"),ID]; colnames(yes)[2]="Yes"
no=casedata[,sum(Attrition=="No"),ID];   colnames(no)[2]="No"
temp=merge(yes,no,all=T)
temp=melt(temp,id=ID)
temp=as.data.frame(temp)
colnames(temp)=c(ID,"Attrition","Count")
temp=merge(temp,as.data.table(temp)[,sum(Count),ID])
colnames(temp)[ncol(temp)]="Total"
temp[["Proportion"]]=temp[["Count"]]/temp[["Total"]]
ggplot(temp,aes(fill=Attrition,y=Proportion,x=eval(as.symbol(ID))))+geom_bar(stat="identity")+xlab(ID)+ggtitle(paste(ID,"vs Attrition"))
```

Here we see attrition decreasing as years with current manager increases with an outlier at year 14 where we only have 5 total observations.
``` {r 14}
temp[temp[["YearsWithCurrManager"]]==14,]

```

