---
title: "MSDS Second Case Study"
author: "Alec Lepe, Che Cobb"
date: "11/28/2018"
output: 
  html_document:
    keep_md: true
---
### Load Libraries
```{r Setup}
#install.packages("data.table")
#install.packages("readxl")
#install.packages("ggplot2")
#install.packages("reshape")
#install.packages("pander")
library("pander")
library("data.table")
library("readxl")
library("ggplot2")
library("reshape")
```
### Reproducibility: Session Info
```{r aeessionInfo}
sessionInfo()
```
### Load Data
```{r LoadData}
casedata=data.table::as.data.table(readxl::read_excel("Data/CaseStudy2-data.xlsx"))
save(casedata,file="Data/CaseStudy2-data.RData")
load("Data/CaseStudy2-data.RData")
dim(casedata)
str(casedata)
```
```{r}
library(corrplot)
library(magrittr)
numeric=casedata %>% dplyr::select(Age,DailyRate,DistanceFromHome,HourlyRate,MonthlyIncome,MonthlyRate,NumCompaniesWorked,PercentSalaryHike,YearsAtCompany,YearsInCurrentRole,YearsSinceLastPromotion,YearsWithCurrManager,TotalWorkingYears,TrainingTimesLastYear,StockOptionLevel)
corrplot(cor(numeric),method="circle",type="upper")
```



### Detect and Delete Non-Relevant Data
```{r DeleteData}
cns=c()
for (cn in colnames(casedata)){
  if(length(table(casedata[[cn]]))==1){
    print(cn)
    cns=append(cn,cns)
  }
}
for (cn in cns){
  print(paste(cn,names(table(casedata[[cn]])),table(casedata[[cn]])))
  casedata=casedata[,!cn,with=F]
}
```
Deleted: *Standard Hours*, *Over18*, and *EmplyoeeCount* where there is only one value.

```{r AttritionCount}
library(dplyr)
library(scales)
casedata %>%group_by(Attrition)%>%summarize(attrition_count=n())%>%ungroup()%>%
  mutate(attrition_rate=round(attrition_count/sum(attrition_count),2))%>%ggplot(aes(x=reorder(Attrition,attrition_rate),y=attrition_rate,fill=Attrition)) + geom_bar(stat='identity',alpha=0.5) +
   theme(axis.text.x=element_text(angle=0,vjust=0.5),legend.position='none',plot.title = element_text(size=12))+labs(title="Attrition Rate")+scale_y_continuous(labels=percent_format())+geom_text(aes(label=attrition_rate),vjust=0.1)

```


### How does age relate to attrition?
```{r AttritionCountByAge}
ID="Age"
yes=casedata[,sum(Attrition=="Yes"),ID]; colnames(yes)[2]="Yes"
no=casedata[,sum(Attrition=="No"),ID];   colnames(no)[2]="No"
temp=merge(yes,no,all=T)
temp=melt(temp,id=ID)
temp=as.data.frame(temp)
colnames(temp)=c(ID,"Attrition","Count")
ggplot(temp,aes(fill=Attrition,y=Count,x=eval(as.symbol(ID))))+geom_bar(stat="identity")+xlab(ID)
```

Here we can see that attrition is greatest between 28 and 35.
```{r AttritionProportionByAge}
temp=merge(temp,as.data.table(temp)[,sum(Count),ID])
colnames(temp)[ncol(temp)]="Total"
temp[["Proportion"]]=temp[["Count"]]/temp[["Total"]]
ggplot(temp,aes(fill=Attrition,y=Proportion,x=eval(as.symbol(ID))))+geom_bar(stat="identity")+xlab(ID)+ggtitle(paste(ID,"vs Attrition"))
```

Here we see that attrition is greater at earlier ages, with a noticeable peak at age 58.
```{r age58}
temp[temp[["Age"]]==58,]
```
### How does business travel relate to attrition?
```{r AttritionCountByBusinessTravel}
ID="BusinessTravel"
yes=casedata[,sum(Attrition=="Yes"),ID]; colnames(yes)[2]="Yes"
no=casedata[,sum(Attrition=="No"),ID];   colnames(no)[2]="No"
temp=merge(yes,no,all=T)
temp=melt(temp,id=ID)
temp=as.data.frame(temp)
colnames(temp)=c(ID,"Attrition","Count")
temp=merge(temp,as.data.table(temp)[,sum(Count),ID])
colnames(temp)[ncol(temp)]="Total"
temp[["Proportion"]]=temp[["Count"]]/temp[["Total"]]
ggplot(temp,aes(fill=Attrition,y=Proportion,x=eval(as.symbol(ID))))+geom_bar(stat="identity")+xlab(ID)+ggtitle(paste(ID,"vs Attrition"))
```

Here we can see that attrition, proportionally, is greatest for employees who travel frequently and the least for employees who do not travel.

### Automatically Detecting Good Single Predictors
We can go on and on plotting individual predictors, but let's try to automate the process.
This R code will go through every numeric variable and see if they can predict attrition.
```{r determinebestvariables}
tempcasedata=casedata
tempcasedata[tempcasedata[["Attrition"]]=="Yes",][["Attrition"]]="1"
tempcasedata[tempcasedata[["Attrition"]]=="No",][["Attrition"]]="0"
mode(tempcasedata[["Attrition"]])="numeric"
pmatrix=matrix(ncol=2,nrow=0)
for (i in setdiff(colnames(tempcasedata),"Attrition")){
  if (mode(tempcasedata[[i]])=='numeric'){
  formula=paste0("Attrition~",i,sep="")
  model=lm(formula,tempcasedata)
  pmatrix=rbind(pmatrix,c(i,summary(model)$coefficients[8]))
}
}
pmatrix=as.data.table(pmatrix[order(as.numeric(pmatrix[,2])),])
colnames(pmatrix)=c("Variable","P-Value")
```
Here are the top 6 variables, based on low P-Values.
```{r variables}
head(pmatrix)
cat(head(pmatrix)[["Variable"]])
```
### Plotting Statistically Significant Data
Excluding age as we previouslyu plotted that out.
```{r TotalWorkingYears}
ID="TotalWorkingYears"
yes=casedata[,sum(Attrition=="Yes"),ID]; colnames(yes)[2]="Yes"
no=casedata[,sum(Attrition=="No"),ID];   colnames(no)[2]="No"
temp=merge(yes,no,all=T)
temp=melt(temp,id=ID)
temp=as.data.frame(temp)
colnames(temp)=c(ID,"Attrition","Count")
temp=merge(temp,as.data.table(temp)[,sum(Count),ID])
colnames(temp)[ncol(temp)]="Total"
temp[["Proportion"]]=temp[["Count"]]/temp[["Total"]]
ggplot(temp,aes(fill=Attrition,y=Proportion,x=eval(as.symbol(ID))))+geom_bar(stat="identity")+xlab(ID)+ggtitle(paste(ID,"vs Attrition"))
```

**We see that attrition generally decreases as total working years increases with an outlier at age 40.**

```{r TotalWorkingYears40}
temp[temp[["TotalWorkingYears"]]==40,]
```
``` {r JobLevel}
ID="JobLevel"
yes=casedata[,sum(Attrition=="Yes"),ID]; colnames(yes)[2]="Yes"
no=casedata[,sum(Attrition=="No"),ID];   colnames(no)[2]="No"
temp=merge(yes,no,all=T)
temp=melt(temp,id=ID)
temp=as.data.frame(temp)
colnames(temp)=c(ID,"Attrition","Count")
temp=merge(temp,as.data.table(temp)[,sum(Count),ID])
colnames(temp)[ncol(temp)]="Total"
temp[["Proportion"]]=temp[["Count"]]/temp[["Total"]]
ggplot(temp,aes(fill=Attrition,y=Proportion,x=eval(as.symbol(ID))))+geom_bar(stat="identity")+xlab(ID)+ggtitle(paste(ID,"vs Attrition"))
```

**Job level varies with attrition**

```{r YearsInCurrentRole}
ID="YearsInCurrentRole"
yes=casedata[,sum(Attrition=="Yes"),ID]; colnames(yes)[2]="Yes"
no=casedata[,sum(Attrition=="No"),ID];   colnames(no)[2]="No"
temp=merge(yes,no,all=T)
temp=melt(temp,id=ID)
temp=as.data.frame(temp)
colnames(temp)=c(ID,"Attrition","Count")
temp=merge(temp,as.data.table(temp)[,sum(Count),ID])
colnames(temp)[ncol(temp)]="Total"
temp[["Proportion"]]=temp[["Count"]]/temp[["Total"]]
ggplot(temp,aes(fill=Attrition,y=Proportion,x=eval(as.symbol(ID))))+geom_bar(stat="identity")+xlab(ID)+ggtitle(paste(ID,"vs Attrition"))
```

**The longer a person is in their current role the longer they stay usually.**



```{r yearsincurrentrole15}
temp[temp[["YearsInCurrentRole"]]==15,]
```

```{r MonthlyIncome}
ID="MonthlyIncome2"
casedata[["MonthlyIncome2"]]=round(casedata[["MonthlyIncome"]]/1000)
yes=casedata[,sum(Attrition=="Yes"),ID]; colnames(yes)[2]="Yes"
no=casedata[,sum(Attrition=="No"),ID];   colnames(no)[2]="No"
temp=merge(yes,no,all=T)
temp=melt(temp,id=ID)
temp=as.data.frame(temp)
colnames(temp)=c(ID,"Attrition","Count")
temp=merge(temp,as.data.table(temp)[,sum(Count),ID])
colnames(temp)[ncol(temp)]="Total"
temp[["Proportion"]]=temp[["Count"]]/temp[["Total"]]
ggplot(temp,aes(fill=Attrition,y=Proportion,x=MonthlyIncome2))+geom_bar(stat="identity")+xlab("Rounded Monthly Income ($K)")+ggtitle("Rounded Monthly Income ($K) vs Attrition")
```

**Attrition generally decreases when monthly income increases.**



```{r YearsWithCurrManager}
ID="YearsWithCurrManager"
yes=casedata[,sum(Attrition=="Yes"),ID]; colnames(yes)[2]="Yes"
no=casedata[,sum(Attrition=="No"),ID];   colnames(no)[2]="No"
temp=merge(yes,no,all=T)
temp=melt(temp,id=ID)
temp=as.data.frame(temp)
colnames(temp)=c(ID,"Attrition","Count")
temp=merge(temp,as.data.table(temp)[,sum(Count),ID])
colnames(temp)[ncol(temp)]="Total"
temp[["Proportion"]]=temp[["Count"]]/temp[["Total"]]
ggplot(temp,aes(fill=Attrition,y=Proportion,x=eval(as.symbol(ID))))+geom_bar(stat="identity")+xlab(ID)+ggtitle(paste(ID,"vs Attrition"))
```

**The longer your usually with your current manager attrition with an outlier at year 14 where there was only have 5 total observations.**


```{r 14}
temp[temp[["YearsWithCurrManager"]]==14,]
```

### Worst Indicators
Now let's take a look at the worst indicators.


```{r WorstIndicators}
tail(pmatrix)
```

```{r worstindicatorplots}
ID="Education"
yes=casedata[,sum(Attrition=="Yes"),ID]; colnames(yes)[2]="Yes"
no=casedata[,sum(Attrition=="No"),ID];   colnames(no)[2]="No"
temp=merge(yes,no,all=T)
temp=melt(temp,id=ID)
temp=as.data.frame(temp)
colnames(temp)=c(ID,"Attrition","Count")
temp=merge(temp,as.data.table(temp)[,sum(Count),ID])
colnames(temp)[ncol(temp)]="Total"
temp[["Proportion"]]=temp[["Count"]]/temp[["Total"]]
ggplot(temp,aes(fill=Attrition,y=Proportion,x=eval(as.symbol(ID))))+geom_bar(stat="identity")+xlab(ID)+ggtitle(paste(ID,"vs Attrition"))

ID="MonthlyRate"
yes=casedata[,sum(Attrition=="Yes"),ID]; colnames(yes)[2]="Yes"
no=casedata[,sum(Attrition=="No"),ID];   colnames(no)[2]="No"
temp=merge(yes,no,all=T)
temp=melt(temp,id=ID)
temp=as.data.frame(temp)
colnames(temp)=c(ID,"Attrition","Count")
temp=merge(temp,as.data.table(temp)[,sum(Count),ID])
colnames(temp)[ncol(temp)]="Total"
temp[["Proportion"]]=temp[["Count"]]/temp[["Total"]]
ggplot(temp,aes(fill=Attrition,y=Proportion,x=eval(as.symbol(ID))))+geom_bar(stat="identity")+xlab(ID)+ggtitle(paste(ID,"vs Attrition"))

ID="PercentSalaryHike"
yes=casedata[,sum(Attrition=="Yes"),ID]; colnames(yes)[2]="Yes"
no=casedata[,sum(Attrition=="No"),ID];   colnames(no)[2]="No"
temp=merge(yes,no,all=T)
temp=melt(temp,id=ID)
temp=as.data.frame(temp)
colnames(temp)=c(ID,"Attrition","Count")
temp=merge(temp,as.data.table(temp)[,sum(Count),ID])
colnames(temp)[ncol(temp)]="Total"
temp[["Proportion"]]=temp[["Count"]]/temp[["Total"]]
ggplot(temp,aes(fill=Attrition,y=Proportion,x=eval(as.symbol(ID))))+geom_bar(stat="identity")+xlab(ID)+ggtitle(paste(ID,"vs Attrition"))

ID="EmployeeNumber"
yes=casedata[,sum(Attrition=="Yes"),ID]; colnames(yes)[2]="Yes"
no=casedata[,sum(Attrition=="No"),ID];   colnames(no)[2]="No"
temp=merge(yes,no,all=T)
temp=melt(temp,id=ID)
temp=as.data.frame(temp)
colnames(temp)=c(ID,"Attrition","Count")
temp=merge(temp,as.data.table(temp)[,sum(Count),ID])
colnames(temp)[ncol(temp)]="Total"
temp[["Proportion"]]=temp[["Count"]]/temp[["Total"]]
ggplot(temp,aes(fill=Attrition,y=Proportion,x=eval(as.symbol(ID))))+geom_bar(stat="identity")+xlab(ID)+ggtitle(paste(ID,"vs Attrition"))

ID="HourlyRate"
yes=casedata[,sum(Attrition=="Yes"),ID]; colnames(yes)[2]="Yes"
no=casedata[,sum(Attrition=="No"),ID];   colnames(no)[2]="No"
temp=merge(yes,no,all=T)
temp=melt(temp,id=ID)
temp=as.data.frame(temp)
colnames(temp)=c(ID,"Attrition","Count")
temp=merge(temp,as.data.table(temp)[,sum(Count),ID])
colnames(temp)[ncol(temp)]="Total"
temp[["Proportion"]]=temp[["Count"]]/temp[["Total"]]
ggplot(temp,aes(fill=Attrition,y=Proportion,x=eval(as.symbol(ID))))+geom_bar(stat="identity")+xlab(ID)+ggtitle(paste(ID,"vs Attrition"))

ID="PerformanceRating"
yes=casedata[,sum(Attrition=="Yes"),ID]; colnames(yes)[2]="Yes"
no=casedata[,sum(Attrition=="No"),ID];   colnames(no)[2]="No"
temp=merge(yes,no,all=T)
temp=melt(temp,id=ID)
temp=as.data.frame(temp)
colnames(temp)=c(ID,"Attrition","Count")
temp=merge(temp,as.data.table(temp)[,sum(Count),ID])
colnames(temp)[ncol(temp)]="Total"
temp[["Proportion"]]=temp[["Count"]]/temp[["Total"]]
ggplot(temp,aes(fill=Attrition,y=Proportion,x=eval(as.symbol(ID))))+geom_bar(stat="identity")+xlab(ID)+ggtitle(paste(ID,"vs Attrition"))
```

**There is no significant trends in the worst rated indicators**

``` {r perfomancetempdb}
temp
```

###Prep Data for logistic modeling
```{r prepdata}
cns=setdiff(colnames(casedata),"Attrition")
for (cn in cns){
  if (mode(casedata[[cn]])!="numeric"){
    casedata[[cn]]=as.factor(casedata[[cn]])
  }
}
casedata[casedata[["Attrition"]]=="Yes",][["Attrition"]]="1"
casedata[casedata[["Attrition"]]=="No",][["Attrition"]]="0"
mode(casedata[["Attrition"]])="numeric"
```

###Select test data set, 70% of given data.
```{r createtesttraindata}
smp<-floor(0.7*nrow(casedata))
set.seed(1337)
ind <- sample(seq_len(nrow(casedata)),size=smp)
train <- casedata[ind,]
test <- casedata[-ind,]
model <- glm(Attrition ~ ., family = 'binomial', data = train)
summary(model)
sort(summary(model)$coefficients[,4])
```

###Select Statistically Significant Variables
alpha = 0.05
```{r selectvariables}
coefs=summary(model)$coefficients
temp=coefs[coefs[,4]<0.05,]
temp[,4][order(temp[,4])]

variables=names(temp[,4][order(temp[,4])])
variables
formula_vars=c()
for (var in variables){
  if (var %in% colnames(casedata)){
    formula_vars=c(formula_vars,var)
  } else {
    letters=strsplit(var,"")[[1]]
    word=c()
    for (letter in letters){
      word=paste0(word,letter,sep="")
      if (word %in% colnames(casedata)){
        formula_vars=c(formula_vars,word)
      }
    }
  }
}
formula_vars=unique(formula_vars)
formula=paste0(paste0(formula_vars,sep="+"),collapse="")
formula=paste0(strsplit(formula,"")[[1]][1:(length(strsplit(formula,"")[[1]])-1)],collapse = "")
formula=paste0("Attrition~",formula,collapse="")
formula
glm(formula,family = 'binomial', data = train)
```

###Run New Linear Model
```{r newmodel}
newmodel <- glm(formula, family = 'binomial', data = train)
pander(newmodel)
sort(summary(newmodel)$coefficients[,4])
```

###Model Evaluation: Training Data Set
```{r trainpredmodel}
predmodel=round(predict(newmodel,train,type='response'),digits=0)
modelcomparison= as.data.table(cbind(predmodel,train$Attrition,train$EmployeeNumber))
colnames(modelcomparison)=c("Predicted","Actual","EmployeeNumber")
modelcomparison
TN=nrow(modelcomparison[modelcomparison[["Predicted"]]==0 & modelcomparison[["Actual"]]==0,])
FN=nrow(modelcomparison[modelcomparison[["Predicted"]]==0 & modelcomparison[["Actual"]]==1,])
TP=nrow(modelcomparison[modelcomparison[["Predicted"]]==1 & modelcomparison[["Actual"]]==1,])
FP=nrow(modelcomparison[modelcomparison[["Predicted"]]==1 & modelcomparison[["Actual"]]==0,])
mat=matrix(ncol=2,nrow=2)
mat[1,1]=TN
mat[1,2]=FP
mat[2,1]=FN
mat[2,2]=TP
colnames(mat)=c("Predicted=0","Predicted=1")
rownames(mat)=c("Actual=0","Actual=1")
print(nrow(train)==sum(mat))
TrainResults=mat
TrainResults
mat=mat/sum(mat)
RelTrainResults=mat
RelTrainResults
cat("Total Training Dataset Accuracy: ",(TP+TN)/sum(TrainResults))
```
###Model Evaluation: Test Data Set
```{r testpredmodel}
predmodel=round(predict(newmodel,test,type='response'),digits=0)
modelcomparison= as.data.table(cbind(predmodel,test$Attrition,test$EmployeeNumber))
colnames(modelcomparison)=c("Predicted","Actual","EmployeeNumber")
modelcomparison
TN=nrow(modelcomparison[modelcomparison[["Predicted"]]==0 & modelcomparison[["Actual"]]==0,])
FN=nrow(modelcomparison[modelcomparison[["Predicted"]]==0 & modelcomparison[["Actual"]]==1,])
TP=nrow(modelcomparison[modelcomparison[["Predicted"]]==1 & modelcomparison[["Actual"]]==1,])
FP=nrow(modelcomparison[modelcomparison[["Predicted"]]==1 & modelcomparison[["Actual"]]==0,])
mat=matrix(ncol=2,nrow=2)
mat[1,1]=TN
mat[1,2]=FP
mat[2,1]=FN
mat[2,2]=TP
colnames(mat)=c("Predicted=0","Predicted=1")
rownames(mat)=c("Actual=0","Actual=1")
print(nrow(test)==sum(mat))
TestResults=mat
TestResults
mat=mat/sum(mat)
RelTestResults=mat
RelTestResults
cat("Total Test Dataset Accuracy: ",(TP+TN)/sum(TestResults))
```

### Forward Variable Selection

``` {r forwardselection1, message=F}
vars=colnames(casedata)
rm(model)
formula="Attrition~"
for (var in vars){
  #print(var)
  if (exists('model')){
    new_formula=paste0(formula,"+",var)
    new_model=glm(new_formula,family="binomial",data=train)
    predmodel=round(predict(new_model,train,type='response'),digits=0)
    modelcomparison= as.data.table(cbind(predmodel,train$Attrition,train$EmployeeNumber))
    colnames(modelcomparison)=c("Predicted","Actual","EmployeeNumber")
    x1=sum(modelcomparison[["Predicted"]]==modelcomparison[["Actual"]])
    predmodel2=round(predict(model,train,type='response'),digits=0)
    modelcomparison2= as.data.table(cbind(predmodel2,train$Attrition,train$EmployeeNumber))
    colnames(modelcomparison2)=c("Predicted","Actual","EmployeeNumber")
    x2=sum(modelcomparison2[["Predicted"]]==modelcomparison2[["Actual"]])
    if (x1>x2){
      formula=new_formula
      model=new_model
      # print(formula)
      # print(x1)
    }
  } else {
    formula=paste0(formula,"+",var)
    model=glm(formula, family = 'binomial', data = train)
  }
}
```
``` {r forwardselection2}
coefs=summary(model)$coefficients
temp=coefs[coefs[,4]<0.05,]
temp[,4][order(temp[,4])]
```
``` {r forwardselection3}
    predmodel=round(predict(model,test,type='response'),digits=0)
    modelcomparison= as.data.table(cbind(predmodel,test$Attrition,test$EmployeeNumber))
    colnames(modelcomparison)=c("Predicted","Actual","EmployeeNumber")
    sum(modelcomparison[["Predicted"]]==modelcomparison[["Actual"]])/nrow(test)
```
``` {r forwardselection4}
TN=nrow(modelcomparison[modelcomparison[["Predicted"]]==0 & modelcomparison[["Actual"]]==0,])
FN=nrow(modelcomparison[modelcomparison[["Predicted"]]==0 & modelcomparison[["Actual"]]==1,])
TP=nrow(modelcomparison[modelcomparison[["Predicted"]]==1 & modelcomparison[["Actual"]]==1,])
FP=nrow(modelcomparison[modelcomparison[["Predicted"]]==1 & modelcomparison[["Actual"]]==0,])
mat=matrix(ncol=2,nrow=2)
mat[1,1]=TN
mat[1,2]=FP
mat[2,1]=FN
mat[2,2]=TP
colnames(mat)=c("Predicted=0","Predicted=1")
rownames(mat)=c("Actual=0","Actual=1")
print(nrow(test)==sum(mat))
mat
mat/sum(mat)
cat("Total Test Dataset Accuracy: ",(TP+TN)/sum(mat))
pander(model)
coefs=summary(model)$coefficients
temp=coefs[coefs[,4]<0.05,]
temp[,4][order(temp[,4])]
```



### Data Analysis & Conclusion
After cleaning the data file. It was determined the following variables were the top single variable parameters that correspond to attrition:

1: TotalWorkingYears

2: JobLevel

3: YearsInCurrentRole

Based on a logistic model on certain variables, with statistical significance of <5%, of a full model here are the top three parameters that correspond to attrition:

1: OverTime

2: EnvironmentSatisfaction 

3: NumCompaniesWorked 

The logistic model also generated a **88% accuracy** on the test dataset.

Based on a forward selection model, here are the top three parameters that correspond to attrition:

1: OverTimeYes

2: YearsSinceLastPromotion

3: YearsInCurrentRole 

The forward variable selection model also generated  a **83% accuracy** on the test dataset.
